\chapter*{Acknowledgements}

First of all I want to thank my main supervisor Jón Egill Kristjánsson for an interesting project and for the opportunity to come and work at NCAR for a couple of weeks. Getting to both work with clouds, which I love, and go to the U.S.A. for the first time makes a happy student. I had a great time over there and am thankful that I could join him and his wife for both hikes and meals. I also want to thank my other supervisor Kari Alterskjær for her "poking" and setting deadlines and pushing me for the past few months -- thank you for being a great motivator! I greatly appreciate the guidance and criticism that they have both given me throughout this project. I have learned a lot from both of them, and hope I get the opportunity to work with them again in the future.

Thank you so much to Anne Claire Fuillioux for helping me with setting up and getting started with WRF, and to both her and Kjell Andresen for help with technical problems during my work with this thesis. Thanks should also be given to Kjetil Schanke Aas for always taking time to answer any stupid WRF-related questions I might have had.

Thanks also to Gregory Thompson whom Jón Egill and I met with in Boulder, for meeting with us and for answering all my e-mails about running the new aerosol-aware microphysics scheme in WRF.

These past two years in the master program at MetOs have been very enjoyable, despite all the stress. Thanks to Marta and Helle especially, for all their positiveness.

I would also like to thank friends and family, for their encouragement, and for rooting for me during my work with this thesis.

Last, but not least, I would like to thank Henrik Andersen Sveinsson, my love and technical helper, for bearing with me through the last few weeks I worked on this thesis.

\begin{flushright}
Oslo, 1st June 2015\\Mari Fenn Kristiansen
\end{flushright}
